---
title: "Presetation"
author: "Weiling Li"
date: "9/19/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(knitr)                ## loading the entire knitr package
library(tidyverse)              ## the only tidyverse library you need is ggplot2
library(esquisse)             ## use esquisse to setup the basic plots

library(kableExtra)
library(magrittr)
opts_chunk$set(echo = TRUE)
```
## Class assignment:

Using the MPG dataset, which is available with the ggplot2 library, produce a document that includes the following elements: headers, text, tables, and plots.  

Tables should include a comparison of city and highway mileage by the class of car and the best three car models for city and highway mileage for all the years in which data is available.  

Plot the data displaying as much of the data as you can. Put continuous variables on the axes and include a locally smoothed regression line to show the relationship with mileage.  Also make a box plot comparing city and highway MPG by class of car.  

## Tables1

```{r,echo = FALSE}
data(mpg)     ## load the data
mpg$cyl <- as.factor(mpg$cyl)   ## convert discrete variables to factors so that they plot
mpg$drv <- as.factor(mpg$drv)   ## as different colors, not gradations of one color
mpg$cty <- as.double(mpg$cty)


mpg_1 <- select(mpg, cty, hwy, class) %>% group_by(class) %>% summarise(City=mean(cty),Highway = mean(hwy)) %>% arrange(class,Highway,City)

kable(mpg_1,format = "html", booktabs=TRUE, digits = 2,     ## call kable to make the table
      col.names = c("Class", "City", "Highway"), 
      caption = "Mean City and Highway MPG by Car Class",align = 'c')%>%
  kable_styling(position = "center",font_size = 14)


```

## Tables2

```{r,echo = FALSE}
years = unique(mpg$year)

mpg_2 <- select(mpg,cty,hwy,model,year) %>% filter(year == years[1]) %>% group_by(model) %>% summarise(City = mean(cty),Highway = mean(hwy)) %>% arrange(model,City,Highway)
mpg_3 <- select(mpg,cty,hwy,model,year) %>% filter(year == years[2]) %>% group_by(model) %>% summarise(City = mean(cty),Highway = mean(hwy)) %>% arrange(model,City,Highway)

cty1999 <- select(mpg_2,model,City)%>%arrange(desc(City))%>%top_n(n = 3,City)
hwy1999 <- select(mpg_2,model,Highway)%>%arrange(desc(Highway))%>%top_n(n = 3,Highway)

cty2008 <- select(mpg_3,model,City)%>%arrange(desc(City))%>%top_n(n = 3,City)

hwy2008 <- select(mpg_3,model,Highway)%>%arrange(desc(Highway))%>%top_n(n = 3,Highway)

tbl_2 <- cbind(cty1999,
               hwy1999,
               cty2008,
               hwy2008)
colnames(tbl_2) <- c('Model', 'Milage', 
                     "Model", "Milage",
                     'Model', 'Milage',
                     "Model", "Milage"
                     )
kable(tbl_2, digits = 2, format = "html", booktabs=TRUE ,caption = "Top 3 MPG Performing Cars: 1999, 2008",align = "c") %>% kable_styling() %>%
  add_header_above(c("City 1999"=2, 
                     "Highway 1999"=2,
                     "City 2008"=2,
                   "Highway 2008"=2))%>%kable_styling(font_size = 12)
#View(mpg_2)
```


## ggplot1

```{r,echo=FALSE}
ggplot(mpg) +
 aes(x = displ, y = cty) +
 geom_point(aes(color=class),size=2,show.legend = F) + geom_smooth(method = 'loess',show.legend = F) +
 scale_color_hue() +
 theme_minimal(base_size = 16) +
 facet_wrap(vars(year)) + 
  labs(x = "Engine Displacement",y = "City Mileage" ,title = "City MPG by Class of Car: 1999, 2008")
``` 

## ggplot2

```{r,echo=FALSE} 
 
 ggplot(mpg) +
 aes(x = displ, y = hwy) +
 geom_point(aes(color=class),size=2,show.legend = F) + geom_smooth(method = 'loess',show.legend = F) +
 scale_color_hue() +
 theme_minimal(base_size = 16) +
 facet_wrap(vars(year)) + 
  labs(x = "Engine Displacement", y = "Highway Mileage", title = "Highway MPG by Class of Car: 1999, 2008")
 
 

##esquisser(mpg)

```

## Boxplot1

```{r,echo=FALSE}
ggplot(mpg) +
 aes(x = class, y = cty) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal(base_size=16) +
  labs(x = "Car Class", y = "City Mileage", title = "City MPG by Class of Car: 1999, 2008")
```
  
## Boxplot2

```{r,echo=FALSE}
ggplot(mpg) +
 aes(x = class, y = hwy) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal(base_size=16) +
  labs(x = "Car Class", y = "Highway Mileage", title = "Highway MPG by Class of Car: 1999, 2008")





```




