---
title: "Regex in R"
author: "MA615"
date: "Fall 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE, echo=TRUE}
library(knitr)
opts_chunk$set(echo = TRUE)

```






## Getting started {.smaller}

Libraries and files for this class

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

library(tidyverse)   ## see https://www.tidyverse.org/

library(gapminder)   ## see https://www.gapminder.org/ and
                     ## https://cran.r-project.org/web/packages/gapminder/index.html

library(magrittr)    ##  see  https://cran.r-project.org/web/packages/magrittr/


```

[rstudio cheatsheets](https://www.rstudio.com/resources/cheatsheets/)


## This class ...

The focus in this class is regex in R.

BUT, keep in mind that regex is used in many places:

-- languages -- editors -- 

The are implemented in most computer languages and in many editors.


## Regular Expressions {.smaller}

Regular expressions describe patterns in strings.

### Regular Expressions in R Cheatsheet

   + Character Classes  
   + Anchors  
   + Quantifiers  

### General Applications

  + Detect patterns  
  + Split strings  
  + Locate patterns  
  + Extract patterns  
  + Replace patterns  

## The Cheatsheet examples ...{.smaller}


```{r}

string <- c("hiphopopotamus", "Rhymenoceros", "time for bottomless lyrics")
pattern <- "t.m"
```

### Detect Patterns

```{r eval = FALSE}

grep(pattern, string)

grep(pattern, string, value=TRUE)

grepl(pattern, string)

str_detect(string, pattern)
```
## The Cheatsheet examples ...{.smaller}

### Split a string using a pattern

```{r eval=FALSE}
strsplit(string, pattern)

str_split(string, pattern)
```

### Locate patterns

```{r eval=FALSE}
regexpr(pattern, string)


gregexpr(pattern, string)

str_locate(string, pattern)

str_locate_all(string, pattern)
```

## The Cheatsheet examples ...{.smaller}

### Extract patterns
```{r eval=FALSE}
regmatches(string, regexpr(pattern, string))

regmatches(string, gregexpr(pattern, string))

str_extract(string, pattern)

str_extract_all(string, pattern)

str_extract_all(string, pattern, simplify=TRUE)

str_match(string, pattern)

str_match_all(string, pattern)
```
## The Cheatsheet examples ...{.smaller}

### Replace patterns
```{r eval=FALSE}
replacement = "BLAH"

sub(pattern, replacement, string)

gsub(pattern, replacement, string)

str_replace(string, pattern, replacement)

str_replace_all(string, pattern, replacement)


```








## Searching for a file {.smaller}
Make a list of the files in the working directory.  
Search for files whose name includes the word "regex"  

```{r echo=TRUE, eval=FALSE}

files <- list.files()
head(files)
grep("regex", files, value=TRUE)
grep("regex", files, value=FALSE)
grep("regex", files, value=TRUE, invert = TRUE)
```

## Now try it with a bigger list

```{r }
files2 <- list.files("c:\\Windows")

writeLines(files2,"file2.txt")

file2a <- read.table("file2.txt")

```

start with grep -- parts of a regular expression  

then variations on grep  





## Now let's generate some random strings 

```{r echo=TRUE}
set.seed(2019) 
word_maker <- function(maxletters, nwords){
  letters = c("A","E","I","O","U","H","K","L","M","N","P","W")
  words = rep(NULL, nwords)
  for(i in 1:nwords){
    words[i] = 
      str_flatten(sample(x = letters, 
                                  size = sample(x=1:maxletters, 
                                                size = 1), 
                                  replace = TRUE))
  }
  return(words)
} 

r_words <- word_maker(maxletters = 10, nwords = 1000)

```

## problems  {.smaller}

### 1 

how many 3 letter words are there?  
how many unique 3 letter words are there?  

### 2 

how many 2 letter words are there?  
how many unique 2 letter words are there?  

### 3

Is the word PAU in the list?  KAI?  


### 4  

HOW MANY WORDS BEGIN WITH "k"?  
how many are unique?



## problem 1 {.smaller}

how many 3 letter words are there?

let_3 <- grep("\\b[[:alpha:]]{3}\\b", r_words, value = TRUE)
let_3a <- grep("\\b[A-Z]{3}\\b", r_words, value=TRUE)

how many unique 3 letter words are there?

let_3u <- unique(let_3)

## problem 2 {.smaller}
how many 2 letter words are there?

let_2 <- grep("\\b[[:alpha:]]{2}\\b", r_words, value = TRUE)

how many unique 2 letter words are there?

let_2u <- unique(let_2)

## problem 3 {.smaller}

Is the word PAU in the list?  KAI?

grep("PAU", r_words)

grep("\\bPAU\\b", r_words)



grep("KAI", r_words)

grep("\\bKAI\\b", r_words)

## Problem 4 {.smaller}


HOW MANY WORDS BEGIN WITH "k"?


k_wrd <- grep("^K", r_words, value=TRUE)

how many are unique?
k_wrd_u <- unique(k_wrd)



## References

[String processing in R book](https://www.gastonsanchez.com/Handling_and_Processing_Strings_in_R.pdf)


[String Processcing in R Blog](http://www.gastonsanchez.com/visually-enforced/resources/how-to/2013/09/22/Handling-and-Processing-Strings-in-R/)



[Regex Tidyverse article](https://stringr.tidyverse.org/articles/regular-expressions.html)

[AWK](https://www.gnu.org/software/gawk/manual/gawk.html)


[hex editor](https://mh-nexus.de/en/hxd/)

[online books](https://bookdown.org/)

[other online resources](https://rmarkdown.rstudio.com/gallery.html)

