---
header-includes:
- \usepackage[fontsize=16pt]{scrextend}
indent: true
output: 
  pdf_document:
    latex_engine: xelatex

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)                ## loading the entire knitr package
library(tidyverse)              ####### change# load tidyverse
library(esquisse)             ## use esquisse to setup the basic plots

library(kableExtra)
library(magrittr)
```

\definecolor{mygray}{gray}{.1}
\fontsize{10}{20}
\setlength{\leftskip}{8pt}
\textcolor{mygray}{Extract From: }

\textcolor{mygray}{Bradley Efron and Trevor Hastie}  

\textcolor{mygray}{ \textit{Computer Age Statistical Inference: Algorithms, Evidence and Data Science}}  

\textcolor{mygray}{ \textit{Cambridge University Press, 2016}}  

*https://web.stanford.edu/~hastie/CASI_files/PDF/casi.pdf*

\setlength{\leftskip}{0pt} 
\fontsize{16}{20}  
\vspace{2.5cm}

Modern Bayesian practice uses various strategies to construct an appropriate "prior" $g(\mu)$ in the absense of prior experience, leaving many statisticians unconvinced by the resulting Bayesian inferences. Our second example illustrates the difficulty.

```{r, include=FALSE}
mechanics <- c(7,44,49,59,34,46,0,32,49,52,44,36,42,5,22,18,41,48,31,42,46,63)
vectors <- c(51,69,41,70,42,40,40,45,57,64,61,59,60,30,57,51,63,38,42,69,49,63)

```