---
title: "Midterm Project"
author: "Weiling Li"
date: "10/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
pacman::p_load(tidyverse,reshape2,plotly,DT,htmlwidgets,kableExtra)
## Self-defined transpose function for tibble
transpose_df <- function(df) {
  t_df <- data.table::transpose(df)
  colnames(t_df) <- rownames(df)
  rownames(t_df) <- colnames(df)
  t_df <- t_df %>%
    tibble::rownames_to_column(.data = .) %>%
    tibble::as_tibble(.)
  return(t_df)
}
```

# World Bank Climate Data

## Load and Clean Data

### Load raw data
In the original dataset, NA value is labeled as "..".
```{r}
ClimateRaw1 = readxl::read_xls("Climate_Change.xls",sheet = 1,na = "..",col_types = "guess")
ClimateRaw2 = readxl::read_xls("Climate_Change.xls",sheet = 2,na = "..",col_types = "guess")
ClimateRaw3 = readxl::read_xls("Climate_Change.xls",sheet = 3,na = "..",col_types = "guess")
```

### Visualize raw data

<iframe src="Raw1.html" height="300" width="900" style="border:none;"></iframe>
<br />
<br />
<iframe src="Raw2.html" height="400" width="900" style="border:none;"></iframe>
<br />
<br />
<iframe src="Raw3.html" height="400" width="900" style="border:none;"></iframe>
<br />

### Cleaning data

When cleaning the datatable, there are several country code which is not actually country. Such as `LIC` represent low income countries and these rows represent an aggregated data. In order to separate these when plotting, we need to save these country codes for later use.
<br />
Pull out region and income codes
```{r}
Regioncode = dplyr::filter(ClimateRaw2,Region == "Aggregates")[-grep("income",ClimateRaw2$`Country name`),] # Both the Income and Region aggregated rows will have a "Aggregates" value in the Region column
Regioncode = Regioncode %>% select(`Country code`) %>% pull()

Incomecode = dplyr::filter(ClimateRaw2,Region == "Aggregates")[grep("income",ClimateRaw2$`Country name`),] 
Incomecode = Incomecode %>% select(`Country code`)%>%pull()
```

There are 57 different variable observed for each country every year from 1990 to 2011(Ofcourse there are NA values,actually quite a lot. some is just missing, and some is measured every 10 years instead of every year.). After carefully examined 3rd sheet in the raw data, 9 variables are kept, they are:
```{r}
Relatedvariables = dplyr::select(ClimateRaw3,`Series code`) %>% 
  slice(1:4,14:15,41:43)%>%
  pull()
print(Relatedvariables)
```


* Total Population `SP.POP.TOTL`
* Population anual growth `SP.POP.GROW`
* GDP `NY.GDP.MKTP.CD`
* GNI per Capita `NY.GNI.PCAP.CD`
* Urban Population `SP.URB.TOTL`
* Urban Population anual growth `SP.URB.GROW`
* CO2 Emission per Capita `EN.ATM.CO2E.PC`
* CO2 Emission per PPP `EN.ATM.CO2E.PP.GD.KD`
* CO2 Total Emission `EN.ATM.CO2E.KT`
<br />

Using the codes to filter original Climate data
```{r}
Climate <-  dplyr::filter(ClimateRaw1,`Series code` %in% Relatedvariables)
print(summary(as.factor(Climate$`Series code`)))
```
<br />
Because the second sheet of the original data contains categorical data indicating a country's income status and geographic info. Assign these to the data table using `dplyr::left_join`, and also spread the table to become longer.

```{r}
Climate <- left_join(Climate,select(ClimateRaw2,`Country code`,`Income group`,Region)) %>%
  melt(measure.vars = 7:28)%>%
  dplyr::rename(year = variable)

kable(Climate%>%dplyr::sample_n(10,replace = FALSE))%>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px")
```
<br />
<br />

Get Series code table, Income code table as well as Region table

```{r}
SeriesCode = ClimateRaw3%>%
  select(`Series code`,`Series name`,Definition)%>%
  dplyr::filter(`Series code` %in% Relatedvariables)

IncomeTable = ClimateRaw2%>%
  select(`Country code`,`Country name`)%>%
  filter(`Country code` %in% Incomecode)

IncomeTable = IncomeTable[c(1,6,5,3,2,4),]# re-order it

RegionTable = ClimateRaw2%>%
  select(`Country code`,`Country name`)%>%
  filter(`Country code` %in% Regioncode)

```

Check NA values in world data `WLD`, Region data and also Income group data

```{r}
Climate%>%dplyr::filter(`Country code`=="WLD",is.na(value))%>%
  dplyr::group_by(year)%>%
  dplyr::select(`Series code`,year,value)%>%
  summary()

Climate%>%dplyr::filter(`Country code`%in% Incomecode,is.na(value))%>%
  dplyr::group_by(year)%>%
  dplyr::select(`Series code`,year,value)%>%
  summary()

Climate%>%dplyr::filter(`Country code`%in% Regioncode,is.na(value))%>%
  dplyr::group_by(year)%>%
  dplyr::select(`Series code`,year,value)%>%
  summary()
```

As shown above, 2011 is missing too many datas, year 2011 will be excluded from the Dataset

```{r}
Climate <- Climate%>%dplyr::filter(year != 2011)
```

check again

```{r}
Climate%>%dplyr::filter(`Country code`=="WLD",is.na(value))%>%
  dplyr::group_by(year)%>%
  dplyr::select(`Series code`,year,value)%>%
  summary()

Climate%>%dplyr::filter(`Country code`%in% Incomecode,is.na(value))%>%
  dplyr::group_by(year)%>%
  dplyr::select(`Series code`,year,value)%>%
  summary()

Climate%>%dplyr::filter(`Country code`%in% Regioncode,is.na(value))%>%
  dplyr::group_by(year)%>%
  dplyr::select(`Series code`,year,value)%>%
  summary()
```

Plot our data to have a taste of what it looks like

```{r}

p <- ggplot(Climate%>%dplyr::filter(`Series code` == Relatedvariables[8],`Country code` %in% Regioncode))+
  aes(x = as.numeric(as.character(year)),y = value,color = `Country name`)+
  geom_line()+xlab("Year")+
  ggtitle(Relatedvariables[8])

plotly_build(p)%>%layout(legend = list(orientation = 'h',x = 0, y = -.3))
```
<br />
<br />
Found North America region is missing. In order to compare regions, we need to reconstruct data for North America.
<br />
```{r}
dplyr::filter(ClimateRaw2,Region != "Aggregates")%>%select(Region)%>%pull()%>%unique()
```

From the dataset 3 countries are labeled in North America region: `USA`, `CAN` and `BMU`. Below are the reconstucting process.

```{r}
# Check if NAS has been occupied
Climate%>%dplyr::filter(`Country code`=="NAS")
# Checked
```

Calculate Total CO2 emission, Total Population, GDP and Urban population

```{r}
# Total CO2 emission
D1 <- Climate%>%
  dplyr::filter(Region == "North America")%>%
  pivot_wider(names_from = year,values_from = value)%>%
  dplyr::filter(`Series code` == Relatedvariables[9])%>%
  dplyr::select(1:8)%>%
  dplyr::slice(1)

D1 <- D1%>%as.data.frame()

D1[1,c(1:2,7:8)] = c("NAS","North America","Aggregates","Aggregates")

D2 = Climate%>%
  dplyr::filter(Region == "North America")%>%
  pivot_wider(names_from = year,values_from = value)%>%
  dplyr::filter(`Series code` == Relatedvariables[9])%>%summarise_at(.vars = 9:29,.funs = sum)
D = dplyr::bind_cols(D1,D2)

# Total Population
D1 <- Climate%>%
  dplyr::filter(Region == "North America")%>%
  pivot_wider(names_from = year,values_from = value)%>%
  dplyr::filter(`Series code` == Relatedvariables[1])%>%
  dplyr::select(1:8)%>%
  dplyr::slice(1)

D1 <- D1%>%as.data.frame()

D1[1,c(1:2,7:8)] = c("NAS","North America","Aggregates","Aggregates")

D2 = Climate%>%
  dplyr::filter(Region == "North America")%>%
  pivot_wider(names_from = year,values_from = value)%>%
  dplyr::filter(`Series code` == Relatedvariables[1])%>%
  summarise_at(.vars = 9:29,.funs = sum)
D <- dplyr::bind_rows(D, dplyr::bind_cols(D1,D2))

# GDP
D1 <- Climate%>%
  dplyr::filter(Region == "North America")%>%
  pivot_wider(names_from = year,values_from = value)%>%
  dplyr::filter(`Series code` == Relatedvariables[3])%>%
  dplyr::select(1:8)%>%
  dplyr::slice(1)
D1 <- D1%>%as.data.frame()

D1[1,c(1:2,7:8)] = c("NAS","North America","Aggregates","Aggregates")

D2 = Climate%>%
  dplyr::filter(Region == "North America")%>%
  pivot_wider(names_from = year,values_from = value)%>%
  dplyr::filter(`Series code` == Relatedvariables[3])%>%
  summarise_at(.vars = 9:29,.funs = sum)

D <- dplyr::bind_rows(D, dplyr::bind_cols(D1,D2))

# Urban Population
D1 <- Climate%>%dplyr::filter(Region == "North America")%>%pivot_wider(names_from = year,values_from = value)%>%dplyr::filter(`Series code` == Relatedvariables[5])%>%dplyr::select(1:8)%>%dplyr::slice(1)
D1 <- D1%>%as.data.frame()
D1[1,c(1:2,7:8)] = c("NAS","North America","Aggregates","Aggregates")
D2 = Climate%>%dplyr::filter(Region == "North America")%>%pivot_wider(names_from = year,values_from = value)%>%dplyr::filter(`Series code` == Relatedvariables[5])%>%summarise_at(.vars = 9:29,.funs = sum)
D <- dplyr::bind_rows(D, dplyr::bind_cols(D1,D2))
```

Calculate CO2 per capita, CO2 per unit PPP
Bermuda's PPP factor is missing from the world bank data, the cloest one I could find is a data given by [federal reserve bank](https://fred.stlouisfed.org/series/PLGDPEBMA670NRUG)

```{r}
pppBer <- 1.98021
pppCan <- 1.21
D_Temp <- D%>%dplyr::select(`Series name`,`1990`:`2010`)%>%
  column_to_rownames(var = "Series name")%>%
  transpose_df()
Temp_Colnames <- colnames(D_Temp)
colnames(D_Temp) <- c("year","CO2","Pop","GDP","Urban")
# Calculate ppp in 2005 for North america
ppp2005 <- Climate%>%dplyr::filter(Region == "North America")%>%
  pivot_wider(names_from = year,values_from = value)%>%
  dplyr::filter(`Series code` == Relatedvariables[3])%>%
  dplyr::select(`Country code`,`2005`)%>%
  column_to_rownames(var = "Country code")%>%
  transpose_df()%>%dplyr::transmute(`North America` = BMU*pppBer + CAN * pppCan + USA)%>%
  pull()

D_Temp = D_Temp%>%dplyr::mutate(CO2percap = CO2*1000/Pop,CO2perGDP = CO2*1000000000/ppp2005)
```

Adjust comparing variables exclude Pop growth and Urban growth and GNI per cap. (For simplicity)

```{r}
SeriesCode <- SeriesCode%>%slice(-2,-4,-6)
# update related Variables
Related_variables <- SeriesCode%>%
  dplyr::select(`Series code`)%>%
  pull()

Temp_Colnames <- SeriesCode%>%
  dplyr::select(`Series name`)%>%
  pull()
#set adequate colname order and rename the D_Temp cols
colnames(D_Temp) <- c("year",Temp_Colnames[c(6,1:5)])
# Transpose back
D_Temp <- D_Temp%>%column_to_rownames(var = "year")%>%transpose_df()
colnames(D_Temp)[colnames(D_Temp)=="rowname"] <- "Series name"
print(colnames(D_Temp))
```

Reconstruct the remaining part of climate data

```{r}
D_ = data.frame()

for (i in 1:length(Related_variables)){
  D1 <- Climate%>%
    dplyr::filter(Region == "North America")%>%
    pivot_wider(names_from = year,values_from = value)%>%
    dplyr::filter(`Series code` == Related_variables[i])%>%
    dplyr::select(1:8)%>%
    dplyr::slice(1)
  D1 <- D1%>%as.data.frame()
  D1[1,c(1:2,7:8)] = c("NAS","North America","Aggregates","Aggregates")
  D_ <- bind_rows(D_,D1)
}
# Combine D_ and D_Temp

D <- left_join(D_,D_Temp,by = "Series name")%>%pivot_longer(cols = 9:29,names_to = "year",values_to = "value")
```

Add these to the original Climate table and we keep the related variables

```{r}
Climate <- bind_rows(Climate,D)
Climate <- Climate%>%
  dplyr::filter(`Series code` %in% Related_variables)
# update Regioncode
Regioncode <- append(Regioncode,"NAS")
```

Replot the previous plot

```{r}
p <- ggplot(Climate%>%dplyr::filter(`Series code` == Related_variables[1],`Country code` %in% Regioncode))+
  aes(x = as.numeric(as.character(year)),y = value,color = `Country name`)+
  geom_line()+xlab("year")+
  ggtitle(Related_variables[1])

plotly_build(p)%>%layout(legend = list(orientation = 'h',x = 0, y = -0.3))
```

Series code is not user friendly, we would use Series name instead

```{r}
SeriesName = SeriesCode%>%dplyr::select(`Series name`)%>%pull()
```

try plot

```{r}
p <- ggplot(Climate%>%dplyr::filter(`Series name` == SeriesName[1],`Country code` %in% c(Incomecode,"WLD")))+
  aes(x = as.numeric(as.character(year)),y = value,color = `Country name`)+
  geom_line()+
  xlab("year")+
  ggtitle(SeriesName[1])

plotly_build(p)%>%layout(legend = list(orientation = 'h',x = 0, y = -0.2))

p <- ggplot(Climate%>%dplyr::filter(`Series name` == SeriesName[2],`Country code` %in% c(Incomecode,"WLD")))+
  aes(x = as.numeric(as.character(year)),y = value,color = `Country name`)+
  geom_line()+
  xlab("year")+
  ggtitle(SeriesName[2])

plotly_build(p)%>%layout(legend = list(orientation = 'h',x = 0, y = -0.2))

p <- ggplot(Climate%>%dplyr::filter(`Series name` == SeriesName[3],`Country code` %in% c(Incomecode,"WLD")))+
  aes(x = as.numeric(as.character(year)),y = value,color = `Country name`)+
  geom_line()+
  xlab("year")+
  ggtitle(SeriesName[3])

plotly_build(p)%>%layout(legend = list(orientation = 'h',x = 0, y = -0.2))

p <- ggplot(Climate%>%dplyr::filter(`Series name` == SeriesName[4],`Country code` %in% c(Incomecode,"WLD")))+
  aes(x = as.numeric(as.character(year)),y = value,color = `Country name`)+
  geom_line()+
  xlab("year")+
  ggtitle(SeriesName[4])

plotly_build(p)%>%layout(legend = list(orientation = 'h',x = 0, y = -0.2))

p <- ggplot(Climate%>%dplyr::filter(`Series name` == SeriesName[5],`Country code` %in% c(Incomecode,"WLD")))+
  aes(x = as.numeric(as.character(year)),y = value,color = `Country name`)+
  geom_line()+
  xlab("year")+
  ggtitle(SeriesName[5])

plotly_build(p)%>%layout(legend = list(orientation = 'h',x = 0, y = -0.2))

p <- ggplot(Climate%>%dplyr::filter(`Series name` == SeriesName[6],`Country code` %in% c(Incomecode,"WLD")))+
  aes(x = as.numeric(as.character(year)),y = value,color = `Country name`)+
  geom_line()+
  xlab("year")+
  ggtitle(SeriesName[6])

plotly_build(p)%>%layout(legend = list(orientation = 'h',x = 0, y = -0.2))
```

We shall see how the data is distributed

```{r}
TempD <- Climate%>%dplyr::filter(`Series name` == SeriesName[1],!`Country code` %in% Incomecode & !`Country code` %in% Regioncode,year %in% c(1992,2008))
p1 <- ggplot(TempD)+
  aes(x =log(value),color = `Region`,fill = `Region`)+facet_wrap(as.factor(TempD$year),nrow = 2)+
  geom_histogram(bins = 30,aes(y = ..density..),position = "dodge",alpha = .2)+
  geom_density(alpha = .3)+ylab("year")+
  xlab(paste0("log ",SeriesName[1]))

plotly_build(p1)#%>%layout(legend = list(orientation = 'h',x = 0, y = -0.2))

TempD <- Climate%>%dplyr::filter(`Series name` == SeriesName[2],!`Country code` %in% Incomecode & !`Country code` %in% Regioncode,year %in% c(1992,2008))
p1 <- ggplot(TempD)+
  aes(x =log(value),color = `Region`,fill = `Region`)+facet_wrap(as.factor(TempD$year),nrow = 2)+
  geom_histogram(bins = 30,aes(y = ..density..),position = "dodge",alpha = .2)+
  geom_density(alpha = .3)+ylab("year")+
  xlab(paste0("log ",SeriesName[2]))

plotly_build(p1)

TempD <- Climate%>%dplyr::filter(`Series name` == SeriesName[3],!`Country code` %in% Incomecode & !`Country code` %in% Regioncode,year %in% c(1992,2008))
p1 <- ggplot(TempD)+
  aes(x =log(value),color = `Region`,fill = `Region`)+facet_wrap(as.factor(TempD$year),nrow = 2)+
  geom_histogram(bins = 30,aes(y = ..density..),position = "dodge",alpha = .2)+
  geom_density(alpha = .3)+ylab("year")+
  xlab(paste0("log ",SeriesName[3]))

plotly_build(p1)

TempD <- Climate%>%dplyr::filter(`Series name` == SeriesName[4],!`Country code` %in% Incomecode & !`Country code` %in% Regioncode,year %in% c(1992,2008))
p1 <- ggplot(TempD)+
  aes(x =log(value),color = `Region`,fill = `Region`)+facet_wrap(as.factor(TempD$year),nrow = 2)+
  geom_histogram(bins = 30,aes(y = ..density..),position = "dodge",alpha = .2)+
  geom_density(alpha = .3)+ylab("year")+
  xlab(paste0("log ",SeriesName[4]))

plotly_build(p1)

TempD <- Climate%>%dplyr::filter(`Series name` == SeriesName[5],!`Country code` %in% Incomecode & !`Country code` %in% Regioncode,year %in% c(1992,2008))
p1 <- ggplot(TempD)+
  aes(x =log(value),color = `Region`,fill = `Region`)+facet_wrap(as.factor(TempD$year),nrow = 2)+
  geom_histogram(bins = 30,aes(y = ..density..),position = "dodge",alpha = .2)+
  geom_density(alpha = .3)+ylab("year")+
  xlab(paste0("log ",SeriesName[5]))

plotly_build(p1)

TempD <- Climate%>%dplyr::filter(`Series name` == SeriesName[6],!`Country code` %in% Incomecode & !`Country code` %in% Regioncode,year %in% c(1992,2008))
p1 <- ggplot(TempD)+
  aes(x =log(value),color = `Region`,fill = `Region`)+facet_wrap(as.factor(TempD$year),nrow = 2)+
  geom_histogram(bins = 30,aes(y = ..density..),position = "dodge",alpha = .2)+
  geom_density(alpha = .3)+ylab("year")+
  xlab(paste0("log ",SeriesName[6]))

plotly_build(p1)
```

try pivot series code,log all input

```{r}
Climate1 <- Climate%>%
  dplyr::select(1:3,7:10)%>%
  pivot_wider(names_from = `Series code`,values_from = value)%>%
  dplyr::mutate_at(.vars = 6:11,.funs = log)

TempD <- Climate1%>%dplyr::filter(!`Country code` %in% Incomecode & !`Country code` %in% Regioncode,year ==2005 )

p3 <- ggplot(TempD)+aes(y = `EN.ATM.CO2E.KT`,x = `SP.POP.TOTL`)+geom_point(aes(size = `NY.GDP.MKTP.CD`,color = `Income group`))+scale_size_continuous(range = c(0,4))+xlab("Total Population, size = GDP")+ylab("Total CO2 Emission")

plotly_build(p3)%>%layout(legend = list(orientation = 'h',x = 0, y = -0.3))

TempD <- Climate1%>%dplyr::filter(!`Country code` %in% Incomecode & !`Country code` %in% Regioncode,year ==2005 )

p3 <- ggplot(TempD)+aes(y = `EN.ATM.CO2E.PC`,x = `SP.POP.TOTL`)+geom_point(aes(size = `NY.GDP.MKTP.CD`,color = `Income group`))+scale_size_continuous(range = c(0,4))+xlab("Total Population, size = GDP")+ylab("Total CO2 Emission per Cap")

plotly_build(p3)%>%layout(legend = list(orientation = 'h',x = 0, y = -0.3))
```

Speculation: Higher GDP, Higher income is related to higher